---
title: "2. Assignment - Mining insults"
author: "Melanija Kraljevska, Jana Štremfelj"
date: "1/7/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Reading data

```{r, include=TRUE, eval=FALSE}

train <- read.table(file = 'insults/train.tsv', sep = '\t', header = TRUE)
test <- read.table(file = 'insults/test.tsv', sep = '\t', header = TRUE)
data <- rbind(train, test)
```

### 1. Cleaning

The function clean_data takes the given data as a parameter and returns a corpus which has punctuation, stopwords and unknown symbols removed. In this case, the parameter data is composed of both train and test files and then splited into training and testing set preserving the same documents as before.

```{r, include=TRUE, eval=FALSE}
clean_data <- function (data) {
  #Remove unknown symbols
  data$text_a <- sapply(data$text_a, function(x) gsub("x[[:alnum:]][[:digit:]]", "", as.character(x)))
  data$text_a <- sapply(data$text_a, function(x) gsub("\\n", "", as.character(x)))
  data$text_a <- sapply(data$text_a, function(x) gsub("\\r", "", as.character(x)))
  data$text_a <- sapply(data$text_a, function(x) gsub("\\t", "", as.character(x)))
  data$text_a <- sapply(data$text_a, function(x) gsub("\\u[[:digit:]][[:digit:]][[:digit:]][[:alnum:]]", "", as.character(x)))
  
  corpus <- Corpus(VectorSource(data$text_a))
  
  #Remove punctuation and stopwords
  corpus <- tm_map(corpus, removeWords, stopwords('english'))
  conn = file("english.stop.txt", open="r")
  mystopwords = readLines(conn)
  close(conn)
  corpus <- tm_map(corpus, removeWords, mystopwords)
  corpus <- tm_map(corpus, removeNumbers)
  corpus <- tm_map(corpus, removePunctuation)
  corpus <- tm_map(corpus, stripWhitespace)
  
  #Anonymize proper nouns
  sent_ann <- Maxent_Sent_Token_Annotator()
  word_ann <- Maxent_Word_Token_Annotator() 
  person_ann <- Maxent_Entity_Annotator(kind = "person")
  location_ann <- Maxent_Entity_Annotator(kind = "location")
  organization_ann <- Maxent_Entity_Annotator(kind = "organization")
  
  entities <- function(annots, kind)
  {
    k <- sapply(annots$features, `[[`, "kind")
    s[annots[k == kind]]
  }
  
  for (i in 1:length(corpus)){
    s <- as.String(content(corpus[[i]]))
    if(nchar(trimws(s)) == 0) next
    
    ann <- annotate(s, list(sent_ann, word_ann, person_ann, location_ann, organization_ann))
    
    corpus <- tm_map(corpus, removeWords, as.vector(entities(ann, "person")))
    corpus <- tm_map(corpus, removeWords, as.vector(entities(ann, "location")))
    corpus <- tm_map(corpus, removeWords, as.vector(entities(ann, "organization")))
    content(corpus[[i]])
  }
  corpus <- tm_map(corpus, stripWhitespace)
  corpus <- tm_map(corpus, content_transformer(tolower))
  corpus <- tm_map(corpus, stemDocument)
  corpus
}

#get corpus for all data
corpus <- clean_data(data)
train_corpus <- corpus[1:dim(train)[1]]
#print first document of train corpus
content(train_corpus[[1]])
test_corpus <- corpus[(dim(train)[1]+1):length(corpus)]
#print first document of test corpus
content(test_corpus[[1]])
```

### 2. Exploration

#### Frequency of words

```{r, include=TRUE, eval=FALSE}
tdm <- TermDocumentMatrix(train_corpus)
termFrequency <- rowSums(as.matrix(tdm))
#termFrequency <- subset(termFrequency, termFrequency >= 10)
v <- sort(rowSums(as.matrix(tdm)),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
qplot(seq(length(termFrequency)),sort(termFrequency), xlab = "index", ylab = "Freq")
head(d, 10)
```

#### Clustering
```{r, include=TRUE, eval=FALSE}
dtm <- DocumentTermMatrix(train_corpus, control = list(weighting=weightTfIdf))
mat <- as.matrix(dtm)
k <- c(2, 4, 8, 16)

for(i in 1:length(k)){
  kmeansResult <- kmeans(mat, k[i])
  tsne.proj <- Rtsne(as.matrix(kmeansResult$cluster), perplexity=10, theta=0.2, dims=2, check_duplicates = F)
  df.tsne <-tsne.proj$Y
  #visualize the cluster assignments
  print(qplot(df.tsne[,1],df.tsne[,2], color = kmeansResult$cluster))
  #plot document representations according to class labels
  print(qplot(df.tsne[,1],df.tsne[,2], color = train$label))
}

```
